---
title: "understand the gender pay gap three ways"
author: "chad allison | 5 december 2022"
output: github_document
---

following [this tutorial](https://juliasilge.com/blog/pay-gap-uk/) from Julia Silge

___

### loading libraries and setting preferences

```{r message = F, warning = F}
library(tidyverse)
library(tidytext)
library(dotwhisker)
library(rsample)

knitr::opts_chunk$set(message = F, warning = F, fig.align = "center")
options(scipen = 999)
theme_set(theme_minimal())
```

### loading data

```{r}
df_raw = read_csv(
  "https://raw.githubusercontent.com/rfordatascience/tidytuesday/master/data/2022/2022-06-28/paygap.csv",
  col_types = cols())

head(df_raw)
```

### counting SIC codes

```{r}
df_raw |>
  select(sic_codes) |>
  separate_rows(sic_codes, sep = ":") |>
  count(sic_codes, sort = T) |>
  head(10)
```

### histogram of median difference in hourly pay

```{r message = F, warning = F}
df_raw |>
  ggplot(aes(diff_median_hourly_percent / 100)) +
  geom_histogram(bins = 25, fill = "springgreen4", alpha = 0.75) +
  scale_x_continuous(limits = c(-0.5, 0.5)) +
  labs(x = "percent difference in median hourly pay") +
  annotate("text", x = -0.275, y = 5000,
           label = "values greater than zero indicate\nmen making more than women")
```

### getting more information on the SIC codes

a .csv file of information on the SIC codes can be downloaded at this link: https://www.gov.uk/government/publications/standard-industrial-classification-of-economic-activities-sic

```{r}
sic_codes = read_csv("sic_codes.csv", col_types = cols()) |>
  janitor::clean_names()

sic_codes
```

### joining SIC codes with original data

```{r}
df_joined = df_raw |>
  select(employer_name, diff_median_hourly_percent, sic_codes) |>
  separate_rows(sic_codes, sep = ":") |>
  left_join(sic_codes, by = c("sic_codes" = "sic_code"))

df_joined
```

### tokenizing the code descriptions

```{r}
df_token = df_joined |>
  unnest_tokens(word, description) |>
  anti_join(get_stopwords(), by = "word") |>
  na.omit()

df_token
```

### filtering to most common words and re-formatting `diff_median_hourly_percent`

```{r}
top_words = df_token |>
  count(word) |>
  filter(!word %in% c("activities", "n.e.c", "general", "non")) |>
  slice_max(n, n = 40) |>
  pull(word)

df = df_token |>
  filter(word %in% top_words) |>
  transmute(diff_wage = diff_median_hourly_percent / 100, word)

df
```

### visual showing increase in men's hourly wages compared to women's for the words in our data

```{r}
df |>
  group_by(word) |>
  summarise(diff_wage = mean(diff_wage)) |>
  mutate(word = fct_reorder(word, diff_wage)) |>
  ggplot(aes(diff_wage, word)) +
  geom_point(alpha = 0.9, size = 2, col = "#B165C7") +
  labs(x = "percent increase in men's hourly wages compared to women's",
       y = NULL)
```

### fitting a linear model

```{r}
paygap_fit = lm(diff_wage ~ word, data = df)
summary(paygap_fit)
```

### visualising model results using `dotwhisker` package

```{r}
tidy(paygap_fit) |>
  mutate(term = str_remove(term, "word"),
         term = fct_reorder(term, -estimate)) %>% # this has to be %>% not |>
  dwplot(vars_order = levels(.$term),
         dot_args = list(size = 2, col = "#598DC8"),
         whisker_args = list(col = "#598DC8")) +
  scale_x_continuous(labels = scales::percent) +
  labs(x = "percent increase in men's hourly wages compared to women's",
       y = NULL)
```

### fitting many models

```{r}
paygap_intervals = reg_intervals(diff_wage ~ word, data = df)

paygap_intervals
```

### visualising modeling results

```{r}
paygap_intervals |>
  mutate(term = str_remove(term, "word"),
         term = fct_reorder(term, .estimate)) |>
  ggplot(aes(.estimate, term)) +
  geom_crossbar(aes(xmin = .lower, xmax = .upper),
                col = "#6C9E72", alpha = 0.75) +
  scale_x_continuous(labels = scales::percent) +
  labs(x = "percent increase in men's hourly wages compared to women's",
       y = NULL)
```



















